---
title: "An Analysis of Billionaires of the World"
author:
  - name: "Akash Agrawal Bejarano, Ama K. Appau, Johnson Odejide, Josh Hall"
    email: {}, {}, bo5104a@american.edu, {}
    affiliation: 
      - name: American University
        city: Washington
        state: DC
        url: www.american.edu
copyright: 
  holder: Machine Teachers
  year: 2023
format: 
  pdf:
    toc: true
    toc-title: Table of Contents
    toc-location: left
    toc-depth: 2
    toc-expand: true
editor: visual
execute: 
  warning: false
  error: false
theme: 
  light: flatly
  dark: darkly
---

## 01. Problem Definition

Here, we define the goals, objectives, and success criteria for your machine learning project as well as determine what type of problem it is (classification, regression, clustering, etc.).

## 02. Data Collection

## 03. EDA

Here, we explore and visualize the data to gain insights, identifying patterns, trends, and potential outliers. We assess the distribution of variables, consider the relationships between variables.

```{r EDA}

```

## 04. Data Preprocessing

Here, we handle missing data and outliers, normalize or standardize data if necessary. We encode categorical variables and split the data into training, validation, and test sets.

```{r DataProcessing}

```

## 05. Feature Engineering

Here, we create relevant features from the raw data. We perform feature selection to identify the most important features, applying domain knowledge to transform data effectively.

```{r FeatureEngineering}

```

## 06. Model Selection

In this section, we choose the appropriate machine learning algorithms for our problem, experiment with different models and techniques to find the best one. We tune hyperparameters for optimal performance.

```{r ModelSelection}

```

## 07. Model Training

Here, we train the selected model on the training data monitor the model performance on the validation set.

```{r ModelTraining}

```

## 08. Model Evaluation

Here, we evaluate the model's performance using appropriate metrics (e.g., accuracy, RMSE). Compare the model's performance to baselines and expectations.

```{r ModelEvaluation}

```

## 09. Model Validation

Here, we use the test data to assess the model's generalization performance to ensure that the model performs well on unseen data.

```{r ModelValidation}

```

## 10. Ethical Considerations

This is not a section perse, it's just for us to ensure that the model and data collection processes adhere to ethical guidelines and do not perpetuate biases or harm.
